---
title: "Assignment"
author:  
- Martijn Koster, 6234119
- Jurrian van de Kraats, 5961688
- Tim Poorthuis, 0651478

date: "2023-02-15"

output:
  bookdown::pdf_book:
   
    toc: false 
    number_sections: true
    citation_package: natbib
    bibliography: references.bib
    keep_tex: true
    fig_caption: true 
    pandoc_args: [
      "--bibliography", "references.bib",
      "--csl", "apa.csl"
    ]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, comment = NA, warning = FALSE,
                      error=FALSE, results='hide',fig.keep='all')
```

\newpage

# Introduction

```{r, include=FALSE}
library(rms)
library(tidyverse)
library(kableExtra)
library(psych)
library(ggdag)
library(dagitty) 
```

```{r, include = FALSE}
data <- read.table("data.txt", sep="\t")
attach(data)
```


In this paper we will examine whether there is a causal effect from elderly people who have received an influenza vaccine and the likelihood of hospitalization. This will be examined by using the patient records dataset and different modeling methods using propensity scores, to account for confounders.

# Methods

## Examening the causal structure

| Domain knowledge was used to create a DAG (Figure \@ref(fig:dag) ) that explains the causal structure present in the data. The variable 'contact with chiropractor (GP)' forms a proxy for the unobserved variable healthy lifestyle. Getting the influenza vaccine is connected to age and contact with GP. Having obtained a influenza vaccination was associated with a lower risk for adverse cardiovascular events \citep{behrouzi}. Having pulmonary disease and diabetes increases the chance for cardiovascular disease \citep{nhg}. Sex forms a confounder in a healthy lifestyle \citep{loef} which in turn causes the afore mentioned diseases. 
| The causal paths of the DAG should be examined to asses whether unbiased causal inference is possible. There is a flow of statistical information through open backdoor paths due to observed and unobserved confounders. The eventual model would need to adjust for these confounders in order to perform an unbiased causal inference.

```{r dag, fig.cap = 'DAG model with Vaccine as exposure and hospitalization as outcome', out.width='50%'}
dag1 <- dagify(
  hlsU ~ sex + oU, # + other unobserved variables that accumulate in this one (like smoking)
  cntct ~ hlsU + age, #healthy life style is unobserved, contact is a proxy
  vacc ~ cntct + age, # if you have not seen your GP in the last 12 months, you are not vaccinated
  pulm ~ hlsU, # rokerslong door niet gezonde levenstijl
  DM ~ age + cntct + hlsU, # diabetes door niet gezonden levenstijl, geen contact met de docter, en leeftijd
  cvd ~ age + cntct + pulm + DM + hlsU+ vacc, # hart en vaatziektes door leeftijd, geen contact met docter, rokerslong, diabetes en niet gezonde levenstijl
  hosp ~ cvd + pulm + DM + cntct + vacc + sex, # hosp door alle enge ziektes en geen griepvaccine. 
  exposure = "vacc", # The treatment
  outcome = "hosp", # The effect of treatment
  
# Give co-ordinates of the variables in the plot
coords = list(x = c(vacc = 1, Age = 5, sex = 9, cvd = 16, pulm = 17, DM = 9, contact = 6, hosp = 25, hlsU = 9, oU = 15),
              y = c(vacc = 8.7, Age = 3, sex = 1, cvd = 13, pulm = 5.7, DM = 9.9, contact = 13, hosp = 8, hlsU = 5, oU = 1))
) 


ggdag_status(dag1) + theme_dag()
```

## Statistical Methods

```{r}
tot_vacc <- sum(vacc)
tot_novacc <- sum(vacc==0)

age_mean <- round(mean(age),2)
age_sd <- round(sd(age),2)


age_vacc_mean <- round(mean(age[vacc==1]),2)
age_novacc_mean <- round(mean(age[vacc==0]),2)
age_vacc_sd <- round(sd(age[vacc==1]),2)
age_novacc_sd <- round(sd(age[vacc==0]),2)

contact_mean <- round(mean(contact),2)
contact_vacc_mean <- round(mean(contact[vacc==1]),2)
contact_novacc_mean <- round(mean(contact[vacc==0]),2)
contact_sd <- round(sd(contact),2)
contact_vacc_sd <- round(sd(contact[vacc==1]),2)
contact_novacc_sd <- round(sd(contact[vacc==0]),2)

pulm_sum <- sum(pulm)
pulm_mean <- round(100*mean(pulm), 2)

pulm_vacc_sum <- sum(pulm[vacc==1])
pulm_vacc_mean <- round(100*mean(pulm[vacc==1]),2)
pulm_novacc_sum <- sum(pulm[vacc==0])
pulm_novacc_mean <- round(100*mean(pulm[vacc==0]),2)

sex_sum <- sum(sex) #1=female
sex_mean <- round(100*mean(sex), 2)

sex_vacc_sum <- sum(sex[vacc==1])
sex_vacc_mean <- round(100*mean(sex[vacc==1]),2)

sex_novacc_sum <- sum(sex[vacc==0])
sex_novacc_mean <- round(100*mean(sex[vacc==0]),2)

cvd_sum <- sum(cvd) #1=female
cvd_mean <- round(100*mean(cvd),2)

cvd_vacc_sum <- sum(cvd[vacc==1])
cvd_vacc_mean <- round(100*mean(cvd[vacc==1]),2)

cvd_novacc_sum <- sum(cvd[vacc==0])
cvd_novacc_mean <- round(100*mean(cvd[vacc==0]),2)

dm_sum <- sum(DM) #1=female
dm_mean <- round(100*mean(DM),2)

dm_vacc_sum <- sum(DM[vacc==1])
dm_vacc_mean <- round(100*mean(DM[vacc==1]),2)

dm_novacc_sum <- sum(DM[vacc==0])
dm_novacc_mean <- round(100*mean(DM[vacc==0]),2)
```

```{r}

age_hosp_mean <- round(mean(age[hosp==1]),2)
age_nohosp_mean <- round(mean(age[hosp==0]),2)
age_hosp_sd <- round(sd(age[hosp==1]),2)
age_nohosp_sd <- round(sd(age[hosp==0]),2)


contact_hosp_mean <- round(mean(contact[hosp==1]),2)
contact_nohosp_mean <- round(mean(contact[hosp==0]),2)

contact_hosp_sd <- round(sd(contact[hosp==1]),2)
contact_nohosp_sd <- round(sd(contact[hosp==0]),2)


pulm_hosp_sum <- sum(pulm[hosp==1])
pulm_hosp_mean <- round(100*mean(pulm[vacc==1]),2)
pulm_nohosp_sum <- sum(pulm[hosp==0])
pulm_nohosp_mean <- round(100*mean(pulm[hosp==0]),2)


sex_hosp_sum <- sum(sex[hosp==1])
sex_hosp_mean <- round(100*mean(sex[hosp==1]),2)

sex_nohosp_sum <- sum(sex[hosp==0])
sex_nohosp_mean <- round(100*mean(sex[hosp==0]),2)


cvd_hosp_sum <- sum(cvd[hosp==1])
cvd_hosp_mean <- round(100*mean(cvd[hosp==1]),2)

cvd_nohosp_sum <- sum(cvd[hosp==0])
cvd_nohosp_mean <- round(100*mean(cvd[hosp==0]),2)


dm_hosp_sum <- sum(DM[hosp==1])
dm_hosp_mean <- round(100*mean(DM[hosp==1]),2)

dm_nohosp_sum <- sum(DM[hosp==0])
dm_nohosp_mean <- round(100*mean(DM[hosp==0]),2)

vacc_hosp_sum <- sum(vacc[hosp==1])
vacc_hosp_mean <- round(100*mean(vacc[hosp==1]),2)

vacc_nohosp_sum <- sum(vacc[hosp==0])
vacc_nohosp_mean <- round(100*mean(vacc[hosp==0]),2)
```

```{r}


characteristics <- c('N', 'Age, mean (SD)', 'Contact, mean (SD)', 'Female, n (%)', 'Pulmonary disease, n (%)', 'Cardiovascular disease, n (%)', 'Diabetes mellitus, n(%)', 'Received Influenza Vaccination, n (%)')
total <- c(nrow(data), paste0(age_mean, ' (', age_sd, ')'),
           paste0(contact_mean, ' (', contact_sd, ')'),
           paste0(sex_sum, ' (', sex_mean, ')'),
           paste0(pulm_sum, ' (', pulm_mean, ')'),
           paste0(cvd_sum, ' (', cvd_mean, ')'),
           paste0(dm_sum, ' (', dm_mean, ')'),
           ""
           )

vaccinated <- c(tot_vacc, paste0(age_vacc_mean, ' (', age_vacc_sd, ')'),
           paste0(contact_vacc_mean, ' (', contact_vacc_sd, ')'),
           paste0(sex_vacc_sum, ' (', sex_vacc_mean, ')'),
           paste0(pulm_vacc_sum, ' (', pulm_vacc_mean, ')'),
           paste0(cvd_vacc_sum, ' (', cvd_vacc_mean, ')'),
           paste0(dm_vacc_sum, ' (', dm_vacc_mean, ')'), 
           ""
           )

not_vaccinated <-  c(tot_novacc, paste0(age_novacc_mean, ' (', age_novacc_sd, ')'),
           paste0(contact_novacc_mean, ' (', contact_novacc_sd, ')'),
           paste0(sex_novacc_sum, ' (', sex_novacc_mean, ')'),
           paste0(pulm_novacc_sum, ' (', pulm_novacc_mean, ')'),
           paste0(cvd_novacc_sum, ' (', cvd_novacc_mean, ')'),
           paste0(dm_novacc_sum, ' (', dm_novacc_mean, ')'),
           ""
           )


hospitalized <- c(sum(hosp), paste0(age_hosp_mean, ' (', age_hosp_sd, ')'),
           paste0(contact_hosp_mean, ' (', contact_hosp_sd, ')'),
           paste0(sex_hosp_sum, ' (', sex_hosp_mean, ')'),
           paste0(pulm_hosp_sum, ' (', pulm_hosp_mean, ')'),
           paste0(cvd_hosp_sum, ' (', cvd_hosp_mean, ')'),
           paste0(dm_hosp_sum, ' (', dm_hosp_mean, ')'), 
           paste0(vacc_hosp_sum, ' (', vacc_hosp_mean, ')')
           )

not_hospitalized <-  c(sum(hosp==0), paste0(age_nohosp_mean, ' (', age_nohosp_sd, ')'),
           paste0(contact_nohosp_mean, ' (', contact_nohosp_sd, ')'),
           paste0(sex_nohosp_sum, ' (', sex_nohosp_mean, ')'),
           paste0(pulm_nohosp_sum, ' (', pulm_nohosp_mean, ')'),
           paste0(cvd_nohosp_sum, ' (', cvd_nohosp_mean, ')'),
           paste0(dm_nohosp_sum, ' (', dm_nohosp_mean, ')'),
           paste0(vacc_nohosp_sum, ' (', vacc_nohosp_mean, ')')
           )



tab1 <- data.frame(
  variables = characteristics, 
  total = total, 
  vaccinated = vaccinated,
  not_vaccinated = not_vaccinated,
  hospitalized,
  not_hospitalized
  
)
```

### Baseline Characteristics

| The electronic patient records dataset consist of eight variables. The variables age and contact with a chiropractor were continues, the variables vaccination status, sex, cardiovascular disease, pulmonary disease, diabetes and hospitalization were binominal. Table \@ref(tab:des) presents the baseline characteristics of `r nrow(data)` individuals who were included in the study.
|   Baseline characteristics stratified by the study outcome indicate that `r sum(hosp)` of the respondents were hospitalized. Respondents who were hospitalized were older (`r age_hosp_mean` vs `r age_nohosp_mean`), had more contact with the GP (`r contact_hosp_mean` vs `r contact_nohosp_mean`), were less often female (`r sex_hosp_mean`%), had more often cardiovascular disease (`r cvd_hosp_mean`% vs `r cvd_nohosp_mean`%) and more often Diabetes (`r dm_hosp_mean`% vs `r dm_nohosp_mean`%).

```{r des, echo = FALSE, results='hold'}
tab1 %>%
  kbl(caption = 'Baseline Characteristics stratified by Influenza vaccination received and Hospitalisation',
      col.names = c('Characteristics', 'Total', 'Yes', 'No', 'Yes', 'No'),
      booktabs = T) %>%
  kable_classic(full_width = F, html_font = 'Cambria') %>% 
  add_header_above(c(" ", " ", "Influenza vaccination received" = 2, 'Hospitalized'=2)) %>% 
  kable_styling(latex_options = "hold_position")

```

### Propensity Scores (PS)

| In order to control for confounding, a PS was estimated by fitting a logistic regression. A PS gives the probability being vaccinated for the respondents. Based on the DAG (Figure \@ref(fig:dag) ) the variables: age, sex, cardiovascular disease, pulmonary disease, diabetis mellitus, and GP contact were used in the PS model. For the variables age and contact a spline is used \citep{Tian}.
|     In Figure \@ref(fig:psscore), the PS for vaccinated and unvaccinated individuals appear to be well-balanced, supporting the positivity assumption, which means that both treatement groups have a chance to get the treatment given the covariates \citep{westreich}.

### PS as Covariate

| The first adjusted model is to use the PS as a covariate. In a observational study the treated and untreated group have an equal distribution given that these are divided in groups of a constant propensity. The propensity score can be used as a baseline variable to account for the dimensional difference between groups since it is assumed that the treatment is unconfounded given this propensity score \citep{schafer}. 

### Inverse Probability Weighting (IPW)

| With the aforementioned propensity scores, IPW is calculated by $\frac{1}{PS}$ for the vaccinated group and $\frac{1}{1-PS}$ for the unvaccinated. IPW creates a pseudo-population by equaling the effect of the confounders mimicking a random control trial. \citep{shiba}.

| Two models were created, one with stabilized- and one with unstabilized weights. When using stabilized weights the numerator model also includes confounders, giving more stable estimates compared to the unstabilized weights \citep{ipw}. In both models bootstrapping was used to account for the inflated sample size of the pseudo population. 

```{r}
ps.formula <- vacc ~ rcs(age) + cvd + DM + rcs(contact) + pulm + sex
ps.fit <- lrm(ps.formula, data=data)
ps.fit$stats['C']
```

```{r, fig.show='hide'}
data$ps.score <- predict(ps.fit,type='fitted')
hist(data$ps.score)
```

```{r psscore, fig.cap = 'Distribution of the Propensity score for participants who received a vaccination compared with those who did not receive a vaccination', out.width='70%' }
data %>% 
  ggplot() + 
  aes(x=ps.score, fill=as.factor(vacc)) + 
  geom_histogram(aes(y=..density..), position='identity', alpha=0.3) + 
  geom_density(aes(color=as.factor(vacc)),alpha=0.1, position='identity') +
  theme_minimal() +
  xlab("PS Score") +
  ylab("Density") +
  labs(fill = "Vaccination") + 
  scale_fill_discrete(labels = c("Unvaccinated", "Vaccinated")) +
  guides(color=FALSE) + xlim(0,1)
```

```{r}
mean(data$ps.score[vacc==0]) ; mean(data$ps.score[vacc==1])
```

```{r}
data$ipw <- with(data, 1 / (ifelse(data$vacc == 1, ps.score, 1-ps.score)))
```

```{r}
num <- glm(vacc ~ 1, family='binomial', data=data)$fitted.values
data$ipw.stab <- data$ipw * num  
```

```{r}
sum(data$ipw.stab)
sum(data$ipw)
nrow(data)

```

# Results

```{r}
crude.fit <- lrm(hosp ~ vacc, data=data)
log.or <- crude.fit$coef[2]
se.log.or <- sqrt(diag(vcov(crude.fit))[2])
res_crude <- round(exp(c(log.or, log.or - 1.96*se.log.or, log.or + 1.96*se.log.or)), 3)


res_crude_c <- round(crude.fit$stats['C'], 3) 
res_crude_p <- round(crude.fit$stats['P'], 3)
```

```{r}
adj1.fit <- lrm(hosp ~ vacc + ps.score, data=data)
log.or <- adj1.fit$coef[2]
se.log.or <- sqrt(diag(vcov(adj1.fit))[2])

res_ps <- round(exp(c(log.or, log.or - 1.96*se.log.or, log.or + 1.96*se.log.or)), 3)

res_ps_c <- round(adj1.fit$stats['C'], 3)
res_ps_p <- round(adj1.fit$stats['P'], 3) 
```

```{r}
adj2.fit <- lrm(hosp ~ vacc, weights = ipw.stab, data=data)
log.or <- adj2.fit$coef[2]
se.log.or <- sqrt(diag(vcov(adj2.fit))[2])
res_ipw <- round(exp(c(log.or, log.or - 1.96*se.log.or, log.or + 1.96*se.log.or)), 3) 


res_ipw_c <- round(adj2.fit$stats['C'], 3)
res_ipw_p <- round(adj2.fit$stats['P'], 3)
```

```{r}
adj3.fit <- lrm(hosp ~ vacc, weights = ipw, data=data)
log.or <- adj3.fit$coef[2]
se.log.or <- sqrt(diag(vcov(adj3.fit))[2])
res_unipw <- round(exp(c(log.or, log.or - 1.96*se.log.or, log.or + 1.96*se.log.or)), 3) 


res_unipw_c <- round(adj2.fit$stats['C'], 3)
res_unipw_p <- round(adj2.fit$stats['P'], 3)
```

```{r}
print_p_value <- function(p_value) {
  if (p_value < 0.001) {
    paste0('<.001')
  } else {
    paste0(p_value)
  }
}
```

```{r}
make_or <- function(OR) {
  paste0(OR[1], ' (', OR[2], ' to ', OR[3],')')
}
```

| Table \@ref(tab:res) shows the crude association between vaccination status and hospitalization was examined using logistic regression, and the odds ratio was found to be `r res_crude[1]` (95% CI: `r res_crude[2]`, `r res_crude[3]`), indicating a non-significant association (p = `r print_p_value(res_crude_p)`). The C-statistic, a measure of discrimination, was `r res_crude_c`, suggesting poor predictive performance of the model.
|     Considering the second model in Table \@ref(tab:res), which presents the PS score as covariate, the Odds Ratio suggests a significant negative association (adjusted OR: `r res_ps[1]`. 95% CI `r res_ps[2]` to `r res_ps[3]`; p `r print_p_value(res_ps_p)`). The C-Statistic of `r res_ps_c` suggests that the model has moderate discriminatory power.
|     The IPW model with unstabilised weights yields significant negative association between the exposure and outcome variable (adjusted OR: `r res_unipw[1]`, 95%CI `r res_unipw[2]` to `r res_unipw[3]`; p `r print_p_value(res_unipw_p)`). The model presents moderate discriminatory power, where C = `r res_unipw_c`.
|     The last model in Table \@ref(tab:res) is an IPW model with stabilised weights. This model also presents a negative association between vaccination and hospitalisation (Adjusted OR: `r res_ipw[1]`, 95% CI `r res_ipw[2]` to `r res_ipw[3]`; p `r print_p_value(res_ipw_p)`). The C-statistic suggests moderate discriminatory power (C= `r res_ipw_c`).

```{r}
models <- c('Unadjusted', 'PS score as covariate', 'Unstabilised IPW', 'Stabilised IPW')

OR_95 <- c(make_or(res_crude), make_or(res_ps), make_or(res_unipw), make_or(res_ipw))

OR_P <- c(print_p_value(res_crude_p), print_p_value(res_ps_p), print_p_value(res_unipw_p), print_p_value(res_ipw_p)) 

OR_C <- c(res_crude_c, res_ps_c, res_unipw_c, res_ipw_c)

tab2 <- data.frame(models,
                      OR_95, 
                      OR_P, 
                      OR_C) 
```

```{r}
n.cat	<- 5	# no. categories to split PS
PS_cat <- ceiling(rank(data$ps.score)*n.cat/length(data$ps.score)) # split PS
for (i in 1:max(PS_cat)){
	print(sapply(split(hosp[PS_cat==i],vacc[PS_cat==i]),mean))} 
```

```{r res, echo = FALSE, results='hold'}
tab2 %>% 
   kbl(caption = 'Association between influenza vaccination and hospitalization (n=40000)',
      col.names = c('Model Specification', 'OR (95% CI)', 'P-Value', 'C-Statistic'),
      booktabs = T) %>%
  kable_classic(full_width = F, html_font = 'Cambria') %>%
  kable_styling(latex_options = "hold_position")
```

# Conclusion / Discussion

-   Conclusions supported by data

-   Other issues (both positive and negative) Maybe write something about matching here?


