---
title: "Incidence"
author: "Martijn Koster, Jurrian van de Kraats, Tim Poorthuis"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    toc: true
    highlight: tango
    theme: paper
    code_download: true
   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, comment = NA, warning = FALSE,
                      error=FALSE, results='hide',fig.keep='all')
```

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(knitr)
library(data.table)
library(tidyverse)
library(readxl)
library(lubridate)
library(DT)
library(epiR)
```

# Assignment {-.tabset}

## Paper {-#pap}

\tableofcontents

```{r}

# load the tables
O_P <- read.csv("Tables/OBSERVATION_PERIODS.csv")
events <- read.csv('Tables/EVENTS.csv')
persons <- read.csv('Tables/PERSONS.csv')
instance <- read.csv('Tables/INSTANCE.csv')
codelist <- read_excel("Tables/CodeList_20210128(1).xlsx", 
    sheet = "ANAPHYL")
codelist2 <- read_excel("Tables/CodeList_20210128(1).xlsx", 
    sheet = "GBS")
```


### Short objective {-#short}

The objective of this report is to analyze the incidence rates of Guillain-Barré Syndrom (gbs) and anaphylactic shock using the PHARMO database, and to present the results in a clear and concise manner.

### Methods {-#meth}

#### PHARMO {-#phar}

The data for this analysis was obtained from the PHARMO database, which is a research institute in the Netherlands that maintains a database of electronic medical records. The Pharmo Data Network is a population-based network of healthcare databases and combines data from different healthcare settings in the Netherlands. These different data sources are linked on a patient level through validated algorithms.



```{r}


dfp1 <- O_P %>% 
  filter(op_start_date <= 20160101, # Remove individuals that do not have at least 1 year of FU before start of study
         op_end_date >= 20170102) # make sure that the individuals at least have 1 day of fu during study period 
dfp1$start_study <- 20170101
```

```{r}
# Make sure that all dates are in Date format YYYY-MM-DD
dfp1 <- dfp1 %>% 
  mutate(op_start_date = as.Date(ymd(op_start_date)),
         op_end_date = as.Date(ymd(op_end_date)),
         start_study = as.Date(ymd(start_study)))
```

```{r}
persons <- persons %>% 
  mutate(birthdate = as.Date(ymd(year_of_birth*10000+month_of_birth*100+day_of_birth)), # calculate the birth date
         age = as.numeric(difftime
                          (as.Date(ymd(20170101)), 
                            birthdate, units='days')) / 365.25, # calculate the age 
         age_bands = cut(age, breaks = c(0,19,39,59,79,Inf), 
                         include.lowest = T,include.highest = T, 
                         labels = c("[0,19]",'[20,39]', '[40,59]', '[60,79]', '[80+]')) # Make sure the ages are divided in age bands 
                         ) %>% 
  filter(age >= 0)
```



#### Study population {-#stupop}

For this research particularly, individuals who were in the study between 01/01/2017 and 31/12/2020 are included. In addition, the study's inclusion and exclusion criteria consist of three requirements. The first requirement is that individuals must have at least one year of follow-up before the study's start date. The second requirement is that individuals must have at least one day of follow-up after the study's start date. Lastly, the study will not consider repeated events, and individuals will be censored at the first occurrence of an event. After fulfilling the aforementioned requirements, we have a study population of _N_ = `r nrow(dfp1)`

#### Data Preperation {-#datprep}

Prior to the analysis, various tables are prepared and then combined in order to obtain the incidence rates. First the OBSERVATION_PERIOD table was loaded and individuals who are not in the above mentioned study population, will be removed. 

After doing this, the PERSONS table was loaded. In this table, an age variable is made with the birth data variable. The age was created with the start of the study period as reference. with this age variable, the following age bands are created: 0-19, 20-39, 40-59, 60-79. The age band 80+ is not created, because the there are no individuals who are older than `r round(max(persons$age))`.  

The next step in the date preparation, was merging the OBSERVATION_PERIODS and PERSON table. We only kept the individuals who are in the OBSERVATION_PERIODS table. 


When the aforementioned tables are merged, Concept sets were created. These sets were created from diagnostic codes from the codesheet. All diagnostic codes and vocabulary of the events used to make incidence rates can be found in Table \@ref(fig:codes)

When the aforementioned tables are merged, Concept sets were created. These sets were created from diagnostic codes from the codesheet. All diagnostic codes and vocabulary of the events we will make incidence rates of can be found in Table \@ref(fig:codes).




```{r codes, results='asis', fig.cap = 'A table'}
anaphyl <- codelist %>% 
  filter(`Coding system`== c('ICD9', 'ICD10/CM'))

gbs <- codelist2 %>% 
  filter(`Coding system`== c('ICD9', 'ICD10/CM'))

codes <- data.table(Diagnosis = paste0('anaphyl_',anaphyl$Algorithm,sep = ""), 
                event_code = anaphyl$Code, 
                Vocabulary = anaphyl$`Coding system`)

gbstemp <- data.table(Diagnosis = paste0('gbs_',gbs$Algorithm,sep = ""), 
                event_code = gbs$Code, 
                Vocabulary = gbs$`Coding system`)

codes <- rbind(codes, gbstemp)

datatable(codes, 
          caption='Diagnostic codes retrieved from the CodeList_20210128 ANAPHYL and GBS sheets ',
          class='compact', 
          rownames=F, 
          colnames=c('Diagnosis', 'Event Code', 'Vocabulary'),
          extensions='Buttons',
          options = list(dom = 'Bfrtip'
                         , buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
          ))## Deze op later moment nog wat mooier maken 

```



Lastly, the EVENTS table is loaded, and merged with the event codes. If the persons had an event before the start of the study, they are removed from the dataset. If a person has more than one event, the second event is removed. 

#### Analyis {-#anal}


For the analysis, incidence rates per 1000 person years for the study population are created per event (i.e. Anaphylactic Shock Narrow, Anaphylactic shock broad, and gbs). The formula for the incidence rates can be found below:


\[
\text{Incidence Rates overall} = \frac{Total \ number \ of \ events}{Total \ person-years} * 1000 
\]

Afterwards, the incidence rates are stratified by year and age-bands. The formulas for these can be found below: 
\[
  \text{Incidence rates by year} = \frac{total \  number \ of \ events \ for \ each \ year} { person-years \ for \ each \ year} * 1000
\]

\[
  \text{Incidence rates by year by age bands} = \frac{Total \ number \ of \ events for \ each \ year \  per \  age \ band}{person-years \ for \ each \ year \  per \ age \ band} * 1000
\]

To assess the technical validation, confidence intervals are used. Exact Poisson confidence limits for the estimated rate are found as the Poisson means, for distributions with the observed number of events and probabilities relevant to the chosen confidence level, divided by time at risk. The relationship between the Poisson and chi-square distributions is employed here (Ulm, 1990):

$$Y_l = \frac{\chi^2_{2Y , a/2}}{2}$$
$$Y_u=\frac{\chi^2_2(Y+1), 1-a/2}{2}$$

where Y is the observed number of events, $Y_l$ and $Y_u$ are lower and upper confidence limits for Y respectively, $\chi^2_{v,a}$ is the chi-square quantile for upper tail probability on ν degrees of freedom.





### Results {-#res}
```{r}
data <- left_join(dfp1, persons, by = "person_id") # merge the OBSERVATION_PERIODS and PERSONS table (keep only individuals who are in OBSERVATIONS_PERIODS table)
```




```{r}
dfp2 <- merge(codes, events, by = 'event_code')
dfp2 <- dfp2 %>% 
  filter(start_date_record > 20170101,
         start_date_record < 20210101) %>% 
  arrange(start_date_record)

dfp2 <- dfp2[!duplicated(dfp2$person_id)]
dfp2$end_date_record <- dfp2$start_date_record
dfp2$is_event <- 1

```

```{r}
dftotal <- left_join(data,dfp2, by = 'person_id')
dftotal$is_event[is.na(dftotal$is_event)] <- 0
```

```{r}
# einde van de studie periode op het moment dat er een event is bij een observatie
dftotal$end_incidence <- as.Date(ymd(dftotal$end_date_record))
# When observation does not have an event, make sure end_incidence is coded as end study period (2021-01-01)
dftotal$end_incidence[is.na(dftotal$end_incidence)] <- as.Date(ymd(20210101)) 
#Calculate how many years each observaiton was in the study. 
dftotal$years_in_study <- as.numeric(difftime
                          (dftotal$end_incidence, 
                            dftotal$start_study, units='days')) / 365.25

```


```{r}
#making IR with 95% CI 
IR_95_CI <- function(events, py, N){
  tmp <- as.matrix(cbind(events, py))
  conf_ints <- round(epi.conf(tmp, ctype = "inc.rate", method = "exact", N = N, design = 1, 
   conf.level = 0.95) * 1000, 2)
  return(conf_ints)
}
```

<!-- Ulm K (1990). A simple method to calculate the confidence interval of a standardized mortality ratio. American Journal of Epidemiology 131: 373 - 375 -->



```{r}
#total PY in study period
TPY <- sum(dftotal$years_in_study)

#calculate the amount of anaph shocks Possible in study period 
TNE_anaphP <- dftotal %>% 
  filter(Diagnosis == 'anaphyl_possible') %>% nrow()

# Incidence Rate for anaph shock possible in study period 


iro_anaphP <-  IR_95_CI(TNE_anaphP, TPY, nrow(data))


#calculate the amount of anaph shocks narrow in study period 
TNE_anaphN <- dftotal %>% 
  filter(Diagnosis == 'anaphyl_narrow') %>% nrow()

# IR for anaph shock narrow in study period
iro_anaphN <- IR_95_CI(TNE_anaphN, TPY, nrow(data))

# amount of gbs cases in study period
TNE_gbs <- dftotal %>% 
  filter(Diagnosis == 'gbs_narrow') %>% nrow()

# IR for gbs in study period
iro_gbs <- IR_95_CI(TNE_gbs, TPY, nrow(data))
```




```{r}
# Add for every observations with an event the date of the event in the dataset. 
dftotal$end_incidence_2017 <- as.Date(ymd(dftotal$end_date_record)) 
# for every observtion who does not have an event, make sure these will get the start of 2018 as end date, since we want to calculate the IR for 2017
dftotal$end_incidence_2017[is.na(dftotal$end_incidence_2017)] <- as.Date(ymd(20180101))

# If observation has an event after 2017, it will become the end of the obs period (2018-01-01)
dftotal$end_incidence_2017 <- as.Date(ifelse(year(dftotal$end_incidence_2017) > 2017, as.Date("2018-01-01"), dftotal$end_incidence_2017), origin = "1970-01-01")
```



```{r}
# Make end incidence for every year
df_2018 <- 
  dftotal %>% 
  filter(!year(end_incidence) == 2017)

df_2019 <- 
  df_2018 %>% 
  filter(!year(end_incidence) == 2018)

df_2020 <- 
  df_2019 %>% 
  filter(!year(end_incidence) == 2019)

df_2021 <- 
  df_2020 %>% 
  filter(!year(end_incidence) == 2020) 



```



```{r}
# calculate for each individual the py observed in 2017. 
dftotal$years_in_study_2017 <- as.numeric(difftime
                          (dftotal$end_incidence_2017, 
                            dftotal$start_study, units='days')) / 365
```


```{r}
# Calculate the events in 2017
events_2017 <- subset(dftotal, year(end_incidence) == 2017) 

#the Total personyears for an event in 2017
TPY_2017 <- sum(dftotal$years_in_study_2017)

# Count the amount of anaphyl possible shocks in 2017 
TNE_anaphP_2017 <- events_2017 %>% 
  filter(Diagnosis == 'anaphyl_possible') %>% nrow()

# calculate the IR in 1000 PY for anaphyl possible in 2017
iro_anaphP_2017 <- IR_95_CI(TNE_anaphP_2017, TPY_2017, nrow(df_total))


# Count the amount of anaphyl shocks narrow in 2017 
TNE_anaphN_2017 <- events_2017 %>% 
  filter(Diagnosis == 'anaphyl_narrow') %>% nrow()

# calculate the IR in 1000 PY for anaphyl narrow in 2017
iro_anaphN_2017 <- IR_95_CI(TNE_anaphN_2017, TPY_2017, nrow(df_total))

# Count the amount of gbs in 2017 
TNE_gbs_2017 <- events_2017 %>% 
  filter(Diagnosis == 'gbs_narrow') %>% nrow()

# calculate the IR in 1000 PY for gbs in 2017
iro_gbs_2017 <- IR_95_CI(TNE_gbs_2017, TPY_2017, nrow(df_total))
```



```{r}
# for every observation who has an event put in the date. 
df_2018$end_incidence_2018 <- as.Date(ymd(df_2018$end_date_record))

#for everybody who does not have an event, make sure the end of incidence is 01-01-2019. 
df_2018$end_incidence_2018[is.na(df_2018$end_incidence_2018)] <- as.Date(ymd(20190101))

# If an observation has the event after 2018, the end of the observation will be 01-01-2019
df_2018$end_incidence_2018 <- as.Date(ifelse(year(df_2018$end_incidence_2018) > 2018, as.Date("2019-01-01"), df_2018$end_incidence_2018), origin = "1970-01-01")
df_2018$start_incidence_2018 <- as.Date(ymd(20180101))

# Count the person years of the observation in 2018
df_2018$years_in_study_2018 <- as.numeric(difftime
                          (df_2018$end_incidence_2018, 
                            df_2018$start_incidence_2018, units='days')) / 365

```


```{r}

# All the events in 2018 in a new dataset
events_2018 <- subset(dftotal, year(end_incidence) == 2018) 

# Total person years observed in 2018
TPY_2018 <- sum(df_2018$years_in_study_2018)

# Amount of events of anaphylatic shock possible in 2018 
TNE_anaphP_2018 <- events_2018 %>% 
  filter(Diagnosis == 'anaphyl_possible') %>% nrow()

# Incidence rate in 1000 PY for anaphalytic shock possible in 2018
iro_anaphP_2018 <- IR_95_CI(TNE_anaphP_2018, TPY_2018, nrow(df_2018))


# Amount of events of anaphylatic shock narrow in 2018 
TNE_anaphN_2018 <- events_2018 %>% 
  filter(Diagnosis == 'anaphyl_narrow') %>% nrow()

# IR in 1000 PY for anaphalytic shock naroow in 2018
iro_anaphN_2018 <-  IR_95_CI(TNE_anaphN_2018, TPY_2018, nrow(df_2018))


# Amount of events of GBS in 2018 
TNE_gbs_2018 <- events_2018 %>% 
  filter(Diagnosis == 'gbs_narrow') %>% nrow()

#IR for gbs in 1000 PY
iro_gbs_2018 <-  IR_95_CI(TNE_gbs_2018, TPY_2018, nrow(df_2018))
```




```{r}
# for every observation who has an event put in the date. 
df_2019$end_incidence_2019 <- as.Date(ymd(df_2019$end_date_record))

#for everybody who does not have an event, make sure the end of incidence is 01-01-2019. 
df_2019$end_incidence_2019[is.na(df_2019$end_incidence_2019)] <- as.Date(ymd(20200101))

# If an observation has the event after 2018, the end of the observation will be 01-01-2019
df_2019$end_incidence_2019 <- as.Date(ifelse(year(df_2019$end_incidence_2019) > 2019, as.Date("2020-01-01"), df_2019$end_incidence_2019), origin = "1970-01-01")
df_2019$start_incidence_2019 <- as.Date(ymd(20190101))

# Count the person years of the observation in 2018
df_2019$years_in_study_2019 <- as.numeric(difftime
                          (df_2019$end_incidence_2019, 
                            df_2019$start_incidence_2019, units='days')) / 365

```


```{r}

# All the events in 2019 in a new dataset
events_2019 <- subset(dftotal, year(end_incidence) == 2019) 

# Total person years observed in 2019
TPY_2019 <- sum(df_2019$years_in_study_2019)

# Amount of events of anaphylatic shock possible in 2019 
TNE_anaphP_2019 <- events_2019 %>% 
  filter(Diagnosis == 'anaphyl_possible') %>% nrow()

# Incidence rate in 1000 PY for anaphalytic shock possible in 2019
iro_anaphP_2019 <-  IR_95_CI(TNE_anaphP_2019, TPY_2019, nrow(df_2019))


# Amount of events of anaphylatic shock narrow in 2019 
TNE_anaphN_2019 <- events_2019 %>% 
  filter(Diagnosis == 'anaphyl_narrow') %>% nrow()

# IR in 1000 PY for anaphalytic shock naroow in 2019
iro_anaphN_2019 <- IR_95_CI(TNE_anaphN_2019, TPY_2019, nrow(df_2019))


# Amount of events of GBS in 2019 
TNE_gbs_2019 <- events_2019 %>% 
  filter(Diagnosis == 'gbs_narrow') %>% nrow()

#IR for gbs in 1000 PY
iro_gbs_2019 <- IR_95_CI(TNE_gbs_2019, TPY_2019, nrow(df_2019))
```


```{r}
# for every observation who has an event put in the date. 
df_2020$end_incidence_2020 <- as.Date(ymd(df_2020$end_date_record))

#for everybody who does not have an event, make sure the end of incidence is 01-01-2021. 
df_2020$end_incidence_2020[is.na(df_2020$end_incidence_2020)] <- as.Date(ymd(20210101))

# If an observation has the event after 2018, the end of the observation will be 01-01-2020
df_2020$end_incidence_2020 <- as.Date(ifelse(year(df_2020$end_incidence_2020) > 2020, as.Date("2021-01-01"), df_2020$end_incidence_2020), origin = "1970-01-01")
df_2020$start_incidence_2020 <- as.Date(ymd(20200101))

# Count the person years of the observation in 2018
df_2020$years_in_study_2020 <- as.numeric(difftime
                          (df_2020$end_incidence_2020, 
                            df_2020$start_incidence_2020, units='days')) / 366

```


```{r}

# All the events in 2020 in a new dataset
events_2020 <- subset(dftotal, year(end_incidence) == 2020) 

# Total person years observed in 2020
TPY_2020 <- sum(df_2020$years_in_study_2020)

# Amount of events of anaphylatic shock possible in 2020 
TNE_anaphP_2020 <- events_2020 %>% 
  filter(Diagnosis == 'anaphyl_possible') %>% nrow()

# Incidence rate in 1000 PY for anaphalytic shock possible in 2020
iro_anaphP_2020 <- IR_95_CI(TNE_anaphP_2020, TPY_2020, nrow(df_2020))


# Amount of events of anaphylatic shock narrow in 2020 
TNE_anaphN_2020 <- events_2020 %>% 
  filter(Diagnosis == 'anaphyl_narrow') %>% nrow()

# IR in 1000 PY for anaphalytic shock naroow in 2020
iro_anaphN_2020 <- IR_95_CI(TNE_anaphN_2020, TPY_2020, nrow(df_2020))


# Amount of events of GBS in 2020 
TNE_gbs_2020 <- events_2020 %>% 
  filter(Diagnosis == 'gbs_narrow') %>% nrow()

#IR for gbs in 1000 PY
iro_gbs_2020 <- IR_95_CI(TNE_gbs_2020, TPY_2020, nrow(df_2020))
```


```{r}
Diagnosis <- c('anaphylaxis_narrow', 'anaphylaxis_narrow', 'anaphylaxis_narrow', 'anaphylaxis_narrow','anaphylaxis_narrow', 'anaphylaxis_possible', 'anaphylaxis_possible', 'anaphylaxis_possible', 'anaphylaxis_possible', 'anaphylaxis_possible', 'gbs_narrow', 'gbs_narrow', 'gbs_narrow', 'gbs_narrow', 'gbs_narrow')
Years <- c('Total', '2017', '2018', '2019', '2020',
           'Total', '2017', '2018', '2019', '2020',
           'Total', '2017', '2018', '2019', '2020')

IR <- c(iro_anaphN, iro_anaphN_2017, iro_anaphN_2018, iro_anaphN_2019, iro_anaphN_2020, iro_anaphP,
        iro_anaphP_2017, iro_anaphP_2018, iro_anaphP_2019, iro_anaphP_2020, iro_gbs, iro_gbs_2017, iro_gbs_2018, iro_gbs_2019, iro_gbs_2020)


tab <- data.frame(Diagnosis = Diagnosis, Year = Years, 'IR per 1000 py' = IR)
```


```{r}

make_year_age_bands <- function(data, age_group) {
  new_data <- data %>% 
    filter(age_bands==age_group) 
  return(new_data)
  }


dftotal0019 <- make_year_age_bands(dftotal, '[0,19]')
dftotal2039 <- make_year_age_bands(dftotal, '[20,39]')
dftotal4059 <- make_year_age_bands(dftotal, '[40,59]')
dftotal6079 <- make_year_age_bands(dftotal, '[60,79]')
dftotal80 <- make_year_age_bands(dftotal, '[80+]')

```


```{r}
# Make function to calculate amount of events
amount_events <- function(data, diagn) {
  events <- data %>% 
    filter(Diagnosis == diagn) %>%  nrow()
  return(events)
}
```


```{r}
#total personyears in study for ageband 0-19
TPY0019 <- sum(dftotal0019$years_in_study)
# amount of events for each event for age_band 0-19
TNE_anaphP0019 <- amount_events(dftotal0019, 'anaphyl_possible')
TNE_anaphN0019 <- amount_events(dftotal0019, 'anaphyl_narrow')
TNE_gbs0019 <- amount_events(dftotal0019, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP0019 <- IR_95_CI(TNE_anaphP0019, TPY0019, nrow(dftotal0019))
iro_anaphN0019 <- IR_95_CI(TNE_anaphN0019, TPY0019, nrow(dftotal0019))
iro_gbs0019 <- IR_95_CI(TNE_gbs0019, TPY0019, nrow(dftotal0019))
```

```{r}
#total personyears in study for ageband 20-39
TPY2039 <- sum(dftotal2039$years_in_study)
# amount of events for each event for age_band 20-39
TNE_anaphP2039 <- amount_events(dftotal2039, 'anaphyl_possible')
TNE_anaphN2039 <- amount_events(dftotal2039, 'anaphyl_narrow')
TNE_gbs2039 <- amount_events(dftotal2039, 'gbs_narrow')

#IR for each event in 1000PY for age_band 20-39 
iro_anaphP2039 <- IR_95_CI(TNE_anaphP2039, TPY2039, nrow(dftotal2039))
iro_anaphN2039 <- IR_95_CI(TNE_anaphN2039, TPY2039, nrow(dftotal2039))
iro_gbs2039 <- IR_95_CI(TNE_gbs2039, TPY2039, nrow(dftotal2039))
```

```{r}
#total personyears in study for ageband 0-19
TPY4059 <- sum(dftotal4059$years_in_study)
# amount of events for each event for age_band 0-19
TNE_anaphP4059 <- amount_events(dftotal4059, 'anaphyl_possible')
TNE_anaphN4059 <- amount_events(dftotal4059, 'anaphyl_narrow')
TNE_gbs4059 <- amount_events(dftotal4059, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP4059 <- IR_95_CI(TNE_anaphP4059, TPY4059, nrow(dftotal4059))
iro_anaphN4059 <- IR_95_CI(TNE_anaphN4059, TPY4059, nrow(dftotal4059))
iro_gbs4059 <- IR_95_CI(TNE_gbs4059, TPY4059, nrow(dftotal4059))
```

```{r}
#total personyears in study for ageband 0-19
TPY6079 <- sum(dftotal6079$years_in_study)
# amount of events for each event for age_band 0-19
TNE_anaphP6079 <- amount_events(dftotal6079, 'anaphyl_possible')
TNE_anaphN6079 <- amount_events(dftotal6079, 'anaphyl_narrow')
TNE_gbs6079 <- amount_events(dftotal6079, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP6079 <- IR_95_CI(TNE_anaphP6079, TPY6079, nrow(dftotal6079))
iro_anaphN6079 <- IR_95_CI(TNE_anaphN6079, TPY6079, nrow(dftotal6079))
iro_gbs6079 <- IR_95_CI(TNE_gbs6079, TPY6079, nrow(dftotal6079))
```



```{r}
#Events in 2017 for age group 00-19
events_2017_0019 <- subset(dftotal0019, year(end_incidence) == 2017) 
#the Total personyears for an event in 2017 age_band0019
TPY_2017_0019 <- sum(dftotal0019$years_in_study_2017)

# Count thenumber of events for age_band 0019 in 2017
TNE_anaphP_0019_2017 <- amount_events(events_2017_0019, 'anaphyl_possible')
TNE_anaphN_0019_2017 <- amount_events(events_2017_0019, 'anaphyl_narrow')
TNE_gbs_0019_2017 <- amount_events(events_2017_0019, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_0019_2017 <- IR_95_CI(TNE_anaphP_0019_2017, TPY_2017_0019, nrow(dftotal0019))
iro_anaphN_0019_2017 <- IR_95_CI(TNE_anaphN_0019_2017, TPY_2017_0019, nrow(dftotal0019))
iro_gbs_0019_2017 <- IR_95_CI(TNE_gbs_0019_2017, TPY_2017_0019, nrow(dftotal0019))
```

```{r}
#Events in 2017 for age group 00-19
events_2017_2039 <- subset(dftotal2039, year(end_incidence) == 2017) 
#the Total personyears for an event in 2017 age_band2039
TPY_2017_2039 <- sum(dftotal2039$years_in_study_2017)

# Count thenumber of events for age_band 2039 in 2017
TNE_anaphP_2039_2017 <- amount_events(events_2017_2039, 'anaphyl_possible')
TNE_anaphN_2039_2017 <- amount_events(events_2017_2039, 'anaphyl_narrow')
TNE_gbs_2039_2017 <- amount_events(events_2017_2039, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_2039_2017 <- IR_95_CI(TNE_anaphP_2039_2017, TPY_2017_2039, nrow(dftotal2039))
iro_anaphN_2039_2017 <- IR_95_CI(TNE_anaphN_2039_2017, TPY_2017_2039, nrow(dftotal2039))
iro_gbs_2039_2017 <- IR_95_CI(TNE_gbs_2039_2017, TPY_2017_2039, nrow(dftotal2039))
```

```{r}
#Events in 2017 for age group 00-19
events_2017_4059 <- subset(dftotal4059, year(end_incidence) == 2017) 
#the Total personyears for an event in 2017 age_band4059
TPY_2017_4059 <- sum(dftotal4059$years_in_study_2017)

# Count thenumber of events for age_band 4059 in 2017
TNE_anaphP_4059_2017 <- amount_events(events_2017_4059, 'anaphyl_possible')
TNE_anaphN_4059_2017 <- amount_events(events_2017_4059, 'anaphyl_narrow')
TNE_gbs_4059_2017 <- amount_events(events_2017_4059, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_4059_2017 <- IR_95_CI(TNE_anaphP_4059_2017, TPY_2017_4059, nrow(dftotal4059))
iro_anaphN_4059_2017 <- IR_95_CI(TNE_anaphN_4059_2017, TPY_2017_4059, nrow(dftotal4059))
iro_gbs_4059_2017 <- IR_95_CI(TNE_gbs_4059_2017, TPY_2017_4059, nrow(dftotal4059))
```

```{r}
#Events in 2017 for age group 00-19
events_2017_6079 <- subset(dftotal6079, year(end_incidence) == 2017) 
#the Total personyears for an event in 2017 age_band6079
TPY_2017_6079 <- sum(dftotal6079$years_in_study_2017)

# Count thenumber of events for age_band 6079 in 2017
TNE_anaphP_6079_2017 <- amount_events(events_2017_6079, 'anaphyl_possible')
TNE_anaphN_6079_2017 <- amount_events(events_2017_6079, 'anaphyl_narrow')
TNE_gbs_6079_2017 <- amount_events(events_2017_6079, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_6079_2017 <- IR_95_CI(TNE_anaphP_6079_2017, TPY_2017_6079, nrow(dftotal6079))
iro_anaphN_6079_2017 <- IR_95_CI(TNE_anaphN_6079_2017, TPY_2017_6079, nrow(dftotal6079))
iro_gbs_6079_2017 <- IR_95_CI(TNE_gbs_6079_2017, TPY_2017_6079, nrow(dftotal6079))
```



```{r}
df_2018_0019 <- make_year_age_bands(df_2018, '[0,19]')
df_2018_2039 <- make_year_age_bands(df_2018, '[20,39]')
df_2018_4059 <- make_year_age_bands(df_2018, '[40,59]')
df_2018_6079 <- make_year_age_bands(df_2018, '[60,79]')
```


```{r}
#Events in 2017 for age group 00-19
events_2018_0019 <- subset(df_2018_0019, year(end_incidence) == 2018) 
#the _2018_ personyears for an event in 2018 age_band0019
TPY_2018_0019 <- sum(df_2018_0019$years_in_study_2018)

# Count thenumber of events for age_band 0019 in 2018
TNE_anaphP_0019_2018 <- amount_events(events_2018_0019, 'anaphyl_possible')
TNE_anaphN_0019_2018 <- amount_events(events_2018_0019, 'anaphyl_narrow')
TNE_gbs_0019_2018 <- amount_events(events_2018_0019, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_0019_2018 <- IR_95_CI(TNE_anaphP_0019_2018, TPY_2018_0019, nrow(df_2018_0019))
iro_anaphN_0019_2018 <- IR_95_CI(TNE_anaphN_0019_2018, TPY_2018_0019, nrow(df_2018_0019))
iro_gbs_0019_2018 <- IR_95_CI(TNE_gbs_0019_2018, TPY_2018_0019, nrow(df_2018_0019))
```

```{r}
#Events in 2018 for age group 00-19
events_2018_2039 <- subset(df_2018_2039, year(end_incidence) == 2018) 
#the _2018_ personyears for an event in 2018 age_band2039
TPY_2018_2039 <- sum(df_2018_2039$years_in_study_2018)

# Count thenumber of events for age_band 2039 in 2018
TNE_anaphP_2039_2018 <- amount_events(events_2018_2039, 'anaphyl_possible')
TNE_anaphN_2039_2018 <- amount_events(events_2018_2039, 'anaphyl_narrow')
TNE_gbs_2039_2018 <- amount_events(events_2018_2039, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_2039_2018 <- IR_95_CI(TNE_anaphP_2039_2018, TPY_2018_2039, nrow(df_2018_2039))
iro_anaphN_2039_2018 <- IR_95_CI(TNE_anaphN_2039_2018, TPY_2018_2039, nrow(df_2018_2039))
iro_gbs_2039_2018 <- IR_95_CI(TNE_gbs_2039_2018, TPY_2018_2039, nrow(df_2018_2039))
```

```{r}
#Events in 2018 for age group 00-19
events_2018_4059 <- subset(df_2018_4059, year(end_incidence) == 2018) 
#the _2018_ personyears for an event in 2018 age_band4059
TPY_2018_4059 <- sum(df_2018_4059$years_in_study_2018)

# Count thenumber of events for age_band 4059 in 2018
TNE_anaphP_4059_2018 <- amount_events(events_2018_4059, 'anaphyl_possible')
TNE_anaphN_4059_2018 <- amount_events(events_2018_4059, 'anaphyl_narrow')
TNE_gbs_4059_2018 <- amount_events(events_2018_4059, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_4059_2018 <- IR_95_CI(TNE_anaphP_4059_2018, TPY_2018_4059, nrow(df_2018_4059))
iro_anaphN_4059_2018 <- IR_95_CI(TNE_anaphN_4059_2018, TPY_2018_4059, nrow(df_2018_4059))
iro_gbs_4059_2018 <- IR_95_CI(TNE_gbs_4059_2018, TPY_2018_4059, nrow(df_2018_4059))
```

```{r}
#Events in 2018 for age group 00-19
events_2018_6079 <- subset(df_2018_6079, year(end_incidence) == 2018) 
#the _2018_ personyears for an event in 2018 age_band6079
TPY_2018_6079 <- sum(df_2018_6079$years_in_study_2018)

# Count thenumber of events for age_band 6079 in 2018
TNE_anaphP_6079_2018 <- amount_events(events_2018_6079, 'anaphyl_possible')
TNE_anaphN_6079_2018 <- amount_events(events_2018_6079, 'anaphyl_narrow')
TNE_gbs_6079_2018 <- amount_events(events_2018_6079, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_6079_2018 <- IR_95_CI(TNE_anaphP_6079_2018, TPY_2018_6079, nrow(df_2018_6079))
iro_anaphN_6079_2018 <- IR_95_CI(TNE_anaphN_6079_2018, TPY_2018_6079, nrow(df_2018_6079))
iro_gbs_6079_2018 <- IR_95_CI(TNE_gbs_6079_2018, TPY_2018_6079, nrow(df_2018_6079))
```



```{r}
df_2019_0019 <- make_year_age_bands(df_2019, '[0,19]')
df_2019_2039 <- make_year_age_bands(df_2019, '[20,39]')
df_2019_4059 <- make_year_age_bands(df_2019, '[40,59]')
df_2019_6079 <- make_year_age_bands(df_2019, '[60,79]')
```


```{r}
#Events in 2017 for age group 00-19
events_2019_0019 <- subset(df_2019_0019, year(end_incidence) == 2019) 
#the _2019_ personyears for an event in 2019 age_band0019
TPY_2019_0019 <- sum(df_2019_0019$years_in_study_2019)

# Count thenumber of events for age_band 0019 in 2019
TNE_anaphP_0019_2019 <- amount_events(events_2019_0019, 'anaphyl_possible')
TNE_anaphN_0019_2019 <- amount_events(events_2019_0019, 'anaphyl_narrow')
TNE_gbs_0019_2019 <- amount_events(events_2019_0019, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_0019_2019 <- IR_95_CI(TNE_anaphP_0019_2019, TPY_2019_0019, nrow(df_2019_0019))
iro_anaphN_0019_2019 <- IR_95_CI(TNE_anaphN_0019_2019, TPY_2019_0019, nrow(df_2019_0019))
iro_gbs_0019_2019 <- IR_95_CI(TNE_gbs_0019_2019, TPY_2019_0019, nrow(df_2019_0019))
```

```{r}
#Events in 2019 for age group 00-19
events_2019_2039 <- subset(df_2019_2039, year(end_incidence) == 2019) 
#the _2019_ personyears for an event in 2019 age_band2039
TPY_2019_2039 <- sum(df_2019_2039$years_in_study_2019)

# Count thenumber of events for age_band 2039 in 2019
TNE_anaphP_2039_2019 <- amount_events(events_2019_2039, 'anaphyl_possible')
TNE_anaphN_2039_2019 <- amount_events(events_2019_2039, 'anaphyl_narrow')
TNE_gbs_2039_2019 <- amount_events(events_2019_2039, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_2039_2019 <- IR_95_CI(TNE_anaphP_2039_2019, TPY_2019_2039, nrow(df_2019_2039))
iro_anaphN_2039_2019 <- IR_95_CI(TNE_anaphN_2039_2019, TPY_2019_2039, nrow(df_2019_2039))
iro_gbs_2039_2019 <- IR_95_CI(TNE_gbs_2039_2019, TPY_2019_2039, nrow(df_2019_2039))
```

```{r}
#Events in 2019 for age group 00-19
events_2019_4059 <- subset(df_2019_4059, year(end_incidence) == 2019) 
#the _2019_ personyears for an event in 2019 age_band4059
TPY_2019_4059 <- sum(df_2019_4059$years_in_study_2019)

# Count thenumber of events for age_band 4059 in 2019
TNE_anaphP_4059_2019 <- amount_events(events_2019_4059, 'anaphyl_possible')
TNE_anaphN_4059_2019 <- amount_events(events_2019_4059, 'anaphyl_narrow')
TNE_gbs_4059_2019 <- amount_events(events_2019_4059, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_4059_2019 <- IR_95_CI(TNE_anaphP_4059_2019, TPY_2019_4059, nrow(df_2019_4059))
iro_anaphN_4059_2019 <- IR_95_CI(TNE_anaphN_4059_2019, TPY_2019_4059, nrow(df_2019_4059))
iro_gbs_4059_2019 <- IR_95_CI(TNE_gbs_4059_2019, TPY_2019_4059, nrow(df_2019_4059))
```

```{r}
#Events in 2019 for age group 00-19
events_2019_6079 <- subset(df_2019_6079, year(end_incidence) == 2019) 
#the _2019_ personyears for an event in 2019 age_band6079
TPY_2019_6079 <- sum(df_2019_6079$years_in_study_2019)

# Count thenumber of events for age_band 6079 in 2019
TNE_anaphP_6079_2019 <- amount_events(events_2019_6079, 'anaphyl_possible')
TNE_anaphN_6079_2019 <- amount_events(events_2019_6079, 'anaphyl_narrow')
TNE_gbs_6079_2019 <- amount_events(events_2019_6079, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_6079_2019 <- IR_95_CI(TNE_anaphP_6079_2019, TPY_2019_6079, nrow(df_2019_6079))
iro_anaphN_6079_2019 <- IR_95_CI(TNE_anaphN_6079_2019, TPY_2019_6079, nrow(df_2019_6079))
iro_gbs_6079_2019 <- IR_95_CI(TNE_gbs_6079_2019, TPY_2019_6079, nrow(df_2019_6079))
```



```{r}
df_2020_0019 <- make_year_age_bands(df_2020, '[0,19]')
df_2020_2039 <- make_year_age_bands(df_2020, '[20,39]')
df_2020_4059 <- make_year_age_bands(df_2020, '[40,59]')
df_2020_6079 <- make_year_age_bands(df_2020, '[60,79]')
```


```{r}
#Events in 2017 for age group 00-19
events_2020_0019 <- subset(df_2020_0019, year(end_incidence) == 2020) 
#the _2020_ personyears for an event in 2020 age_band0019
TPY_2020_0019 <- sum(df_2020_0019$years_in_study_2020)

# Count thenumber of events for age_band 0019 in 2020
TNE_anaphP_0019_2020 <- amount_events(events_2020_0019, 'anaphyl_possible')
TNE_anaphN_0019_2020 <- amount_events(events_2020_0019, 'anaphyl_narrow')
TNE_gbs_0019_2020 <- amount_events(events_2020_0019, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_0019_2020 <- IR_95_CI(TNE_anaphP_0019_2020, TPY_2020_0019, nrow(df_2020_0019))
iro_anaphN_0019_2020 <- IR_95_CI(TNE_anaphN_0019_2020, TPY_2020_0019, nrow(df_2020_0019))
iro_gbs_0019_2020 <- IR_95_CI(TNE_gbs_0019_2020, TPY_2020_0019, nrow(df_2020_0019))
```

```{r}
#Events in 2020 for age group 00-19
events_2020_2039 <- subset(df_2020_2039, year(end_incidence) == 2020) 
#the _2020_ personyears for an event in 2020 age_band2039
TPY_2020_2039 <- sum(df_2020_2039$years_in_study_2020)

# Count thenumber of events for age_band 2039 in 2020
TNE_anaphP_2039_2020 <- amount_events(events_2020_2039, 'anaphyl_possible')
TNE_anaphN_2039_2020 <- amount_events(events_2020_2039, 'anaphyl_narrow')
TNE_gbs_2039_2020 <- amount_events(events_2020_2039, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_2039_2020 <- IR_95_CI(TNE_anaphP_2039_2020, TPY_2020_2039, nrow(df_2020_2039))
iro_anaphN_2039_2020 <- IR_95_CI(TNE_anaphN_2039_2020, TPY_2020_2039, nrow(df_2020_2039))
iro_gbs_2039_2020 <- IR_95_CI(TNE_gbs_2039_2020, TPY_2020_2039, nrow(df_2020_2039))
```

```{r}
#Events in 2020 for age group 00-19
events_2020_4059 <- subset(df_2020_4059, year(end_incidence) == 2020) 
#the _2020_ personyears for an event in 2020 age_band4059
TPY_2020_4059 <- sum(df_2020_4059$years_in_study_2020)

# Count thenumber of events for age_band 4059 in 2020
TNE_anaphP_4059_2020 <- amount_events(events_2020_4059, 'anaphyl_possible')
TNE_anaphN_4059_2020 <- amount_events(events_2020_4059, 'anaphyl_narrow')
TNE_gbs_4059_2020 <- amount_events(events_2020_4059, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_4059_2020 <- IR_95_CI(TNE_anaphP_4059_2020, TPY_2020_4059, nrow(df_2020_4059))
iro_anaphN_4059_2020 <- IR_95_CI(TNE_anaphN_4059_2020, TPY_2020_4059, nrow(df_2020_4059))
iro_gbs_4059_2020 <- IR_95_CI(TNE_gbs_4059_2020, TPY_2020_4059, nrow(df_2020_4059))
```

```{r}
#Events in 2020 for age group 00-19
events_2020_6079 <- subset(df_2020_6079, year(end_incidence) == 2020) 
#the _2020_ personyears for an event in 2020 age_band6079
TPY_2020_6079 <- sum(df_2020_6079$years_in_study_2020)

# Count thenumber of events for age_band 6079 in 2020
TNE_anaphP_6079_2020 <- amount_events(events_2020_6079, 'anaphyl_possible')
TNE_anaphN_6079_2020 <- amount_events(events_2020_6079, 'anaphyl_narrow')
TNE_gbs_6079_2020 <- amount_events(events_2020_6079, 'gbs_narrow')

#IR for each event in 1000PY for age_band 0-19 
iro_anaphP_6079_2020 <- IR_95_CI(TNE_anaphP_6079_2020, TPY_2020_6079, nrow(df_2020_6079))
iro_anaphN_6079_2020 <- IR_95_CI(TNE_anaphN_6079_2020, TPY_2020_6079, nrow(df_2020_6079))
iro_gbs_6079_2020 <- IR_95_CI(TNE_gbs_6079_2020, TPY_2020_6079, nrow(df_2020_6079))
```



```{r}
Diag1 <- rep('Anaphylaxis narrow', times=25)
Diag2 <- rep('Anaphylaxis possible', times=25)
Diag3 <- rep('gbs narrow', times=25)
all_diag <- c(Diag1, Diag2, Diag3)

all_years <- rep(c('Total', '2017', '2018', '2019', '2020'), times=15)

rep_total <- rep('Total', times=5)
rep_0019 <- rep('0-19', times=5)
rep_2039 <- rep('20-39', times=5)
rep_4059 <- rep('40-59', times=5)
rep_6079 <- rep('60-79', times=5)

all_ages <- rep(c(rep_total, rep_0019, rep_2039, rep_4059, rep_6079), times=3)



count_events <- c(
  TNE_anaphN, TNE_anaphN_2017, TNE_anaphN_2018, 
  TNE_anaphN_2019, TNE_anaphN_2020, 
  TNE_anaphN0019, TNE_anaphN_0019_2017, TNE_anaphN_0019_2018,
  TNE_anaphN_0019_2019, TNE_anaphN_0019_2020, 
  TNE_anaphN2039, TNE_anaphN_2039_2017, TNE_anaphN_2039_2018,
  TNE_anaphN_2039_2019, TNE_anaphN_2039_2020,
  TNE_anaphN4059, TNE_anaphN_4059_2017, TNE_anaphN_4059_2018,
  TNE_anaphN_4059_2019, TNE_anaphN_4059_2020,
  TNE_anaphN6079, TNE_anaphN_6079_2017, TNE_anaphN_6079_2018,
  TNE_anaphN_6079_2019, TNE_anaphN_6079_2020,
  #### Anaph Possible 
  TNE_anaphP, TNE_anaphP_2017, TNE_anaphP_2018, 
  TNE_anaphP_2019, TNE_anaphP_2020, 
  TNE_anaphP0019, TNE_anaphP_0019_2017, TNE_anaphP_0019_2018,
  TNE_anaphP_0019_2019, TNE_anaphP_0019_2020, 
  TNE_anaphP2039, TNE_anaphP_2039_2017, TNE_anaphP_2039_2018,
  TNE_anaphP_2039_2019, TNE_anaphP_2039_2020,
  TNE_anaphP4059, TNE_anaphP_4059_2017, TNE_anaphP_4059_2018,
  TNE_anaphP_4059_2019, TNE_anaphP_4059_2020,
  TNE_anaphP6079, TNE_anaphP_6079_2017, TNE_anaphP_6079_2018,
  TNE_anaphP_6079_2019, TNE_anaphP_6079_2020,
  #### GBS 
  TNE_gbs, TNE_gbs_2017, TNE_gbs_2018, 
  TNE_gbs_2019, TNE_gbs_2020, 
  TNE_gbs0019, TNE_gbs_0019_2017, TNE_gbs_0019_2018,
  TNE_gbs_0019_2019, TNE_gbs_0019_2020, 
  TNE_gbs2039, TNE_gbs_2039_2017, TNE_gbs_2039_2018,
  TNE_gbs_2039_2019, TNE_gbs_2039_2020,
  TNE_gbs4059, TNE_gbs_4059_2017, TNE_gbs_4059_2018,
  TNE_gbs_4059_2019, TNE_gbs_4059_2020,
  TNE_gbs6079, TNE_gbs_6079_2017, TNE_gbs_6079_2018,
  TNE_gbs_6079_2019, TNE_gbs_6079_2020
  )


PersYear <- c(TPY, TPY_2017, TPY_2018, 
  TPY_2019, TPY_2020, 
  TPY0019, TPY_2017_0019, TPY_2018_0019,
  TPY_2019_0019, TPY_2020_0019, 
  TPY2039, TPY_2017_2039, TPY_2018_2039,
  TPY_2019_2039, TPY_2020_2039,
  TPY4059, TPY_2017_4059, TPY_2018_4059,
  TPY_2019_4059, TPY_2020_4059,
  TPY6079, TPY_2017_4059, TPY_2018_4059,
  TPY_2019_4059, TPY_2020_4059)


pers_year_all <- rep(PersYear, times=3)

```

```{r}
IR_1000PY <-  as.numeric(c(
  iro_anaphN[1], iro_anaphN_2017[1], iro_anaphN_2018[1], 
  iro_anaphN_2019[1], iro_anaphN_2020[1], 
  iro_anaphN0019[1], iro_anaphN_0019_2017[1], iro_anaphN_0019_2018[1],
  iro_anaphN_0019_2019[1], iro_anaphN_0019_2020[1], 
  iro_anaphN2039[1], iro_anaphN_2039_2017[1], iro_anaphN_2039_2018[1],
  iro_anaphN_2039_2019[1], iro_anaphN_2039_2020[1],
  iro_anaphN4059[1], iro_anaphN_4059_2017[1], iro_anaphN_4059_2018[1],
  iro_anaphN_4059_2019[1], iro_anaphN_4059_2020[1],
  iro_anaphN6079[1], iro_anaphN_6079_2017[1], iro_anaphN_6079_2018[1],
  iro_anaphN_6079_2019[1], iro_anaphN_6079_2020[1],
  #### Anaph Possible 
  iro_anaphP[1], iro_anaphP_2017[1], iro_anaphP_2018[1], 
  iro_anaphP_2019[1], iro_anaphP_2020[1], 
  iro_anaphP0019[1], iro_anaphP_0019_2017[1], iro_anaphP_0019_2018[1],
  iro_anaphP_0019_2019[1], iro_anaphP_0019_2020[1], 
  iro_anaphP2039[1], iro_anaphP_2039_2017[1], iro_anaphP_2039_2018[1],
  iro_anaphP_2039_2019[1], iro_anaphP_2039_2020[1],
  iro_anaphP4059[1], iro_anaphP_4059_2017[1], iro_anaphP_4059_2018[1],
  iro_anaphP_4059_2019[1], iro_anaphP_4059_2020[1],
  iro_anaphP6079[1], iro_anaphP_6079_2017[1], iro_anaphP_6079_2018[1],
  iro_anaphP_6079_2019[1], iro_anaphP_6079_2020[1],
  #### GBS 
    iro_gbs[1], iro_gbs_2017[1], iro_gbs_2018[1], 
  iro_gbs_2019[1], iro_gbs_2020[1], 
  iro_gbs0019[1], iro_gbs_0019_2017[1], iro_gbs_0019_2018[1],
  iro_gbs_0019_2019[1], iro_gbs_0019_2020[1], 
  iro_gbs2039[1], iro_gbs_2039_2017[1], iro_gbs_2039_2018[1],
  iro_gbs_2039_2019[1], iro_gbs_2039_2020[1],
  iro_gbs4059[1], iro_gbs_4059_2017[1], iro_gbs_4059_2018[1],
  iro_gbs_4059_2019[1], iro_gbs_4059_2020[1],
  iro_gbs6079[1], iro_gbs_6079_2017[1], iro_gbs_6079_2018[1],
  iro_gbs_6079_2019[1], iro_gbs_6079_2020[1]
  ))
```

```{r}
IR_1000PY_95L <- as.numeric(c(
  iro_anaphN[2], iro_anaphN_2017[2], iro_anaphN_2018[2], 
  iro_anaphN_2019[2], iro_anaphN_2020[2], 
  iro_anaphN0019[2], iro_anaphN_0019_2017[2], iro_anaphN_0019_2018[2],
  iro_anaphN_0019_2019[2], iro_anaphN_0019_2020[2], 
  iro_anaphN2039[2], iro_anaphN_2039_2017[2], iro_anaphN_2039_2018[2],
  iro_anaphN_2039_2019[2], iro_anaphN_2039_2020[2],
  iro_anaphN4059[2], iro_anaphN_4059_2017[2], iro_anaphN_4059_2018[2],
  iro_anaphN_4059_2019[2], iro_anaphN_4059_2020[2],
  iro_anaphN6079[2], iro_anaphN_6079_2017[2], iro_anaphN_6079_2018[2],
  iro_anaphN_6079_2019[2], iro_anaphN_6079_2020[2],
  #### Anaph Possible 
  iro_anaphP[2], iro_anaphP_2017[2], iro_anaphP_2018[2], 
  iro_anaphP_2019[2], iro_anaphP_2020[2], 
  iro_anaphP0019[2], iro_anaphP_0019_2017[2], iro_anaphP_0019_2018[2],
  iro_anaphP_0019_2019[2], iro_anaphP_0019_2020[2], 
  iro_anaphP2039[2], iro_anaphP_2039_2017[2], iro_anaphP_2039_2018[2],
  iro_anaphP_2039_2019[2], iro_anaphP_2039_2020[2],
  iro_anaphP4059[2], iro_anaphP_4059_2017[2], iro_anaphP_4059_2018[2],
  iro_anaphP_4059_2019[2], iro_anaphP_4059_2020[2],
  iro_anaphP6079[2], iro_anaphP_6079_2017[2], iro_anaphP_6079_2018[2],
  iro_anaphP_6079_2019[2], iro_anaphP_6079_2020[2],
  #### GBS 
    iro_gbs[2], iro_gbs_2017[2], iro_gbs_2018[2], 
  iro_gbs_2019[2], iro_gbs_2020[2], 
  iro_gbs0019[2], iro_gbs_0019_2017[2], iro_gbs_0019_2018[2],
  iro_gbs_0019_2019[2], iro_gbs_0019_2020[2], 
  iro_gbs2039[2], iro_gbs_2039_2017[2], iro_gbs_2039_2018[2],
  iro_gbs_2039_2019[2], iro_gbs_2039_2020[2],
  iro_gbs4059[2], iro_gbs_4059_2017[2], iro_gbs_4059_2018[2],
  iro_gbs_4059_2019[2], iro_gbs_4059_2020[2],
  iro_gbs6079[2], iro_gbs_6079_2017[2], iro_gbs_6079_2018[2],
  iro_gbs_6079_2019[2], iro_gbs_6079_2020[2]
  ))
```


```{r}
IR_1000PY_95H <- as.numeric(c(
  iro_anaphN[3], iro_anaphN_2017[3], iro_anaphN_2018[3], 
  iro_anaphN_2019[3], iro_anaphN_2020[3], 
  iro_anaphN0019[3], iro_anaphN_0019_2017[3], iro_anaphN_0019_2018[3],
  iro_anaphN_0019_2019[3], iro_anaphN_0019_2020[3], 
  iro_anaphN2039[3], iro_anaphN_2039_2017[3], iro_anaphN_2039_2018[3],
  iro_anaphN_2039_2019[3], iro_anaphN_2039_2020[3],
  iro_anaphN4059[3], iro_anaphN_4059_2017[3], iro_anaphN_4059_2018[3],
  iro_anaphN_4059_2019[3], iro_anaphN_4059_2020[3],
  iro_anaphN6079[3], iro_anaphN_6079_2017[3], iro_anaphN_6079_2018[3],
  iro_anaphN_6079_2019[3], iro_anaphN_6079_2020[3],
  #### Anaph Possible 
  iro_anaphP[3], iro_anaphP_2017[3], iro_anaphP_2018[3], 
  iro_anaphP_2019[3], iro_anaphP_2020[3], 
  iro_anaphP0019[3], iro_anaphP_0019_2017[3], iro_anaphP_0019_2018[3],
  iro_anaphP_0019_2019[3], iro_anaphP_0019_2020[3], 
  iro_anaphP2039[3], iro_anaphP_2039_2017[3], iro_anaphP_2039_2018[3],
  iro_anaphP_2039_2019[3], iro_anaphP_2039_2020[3],
  iro_anaphP4059[3], iro_anaphP_4059_2017[3], iro_anaphP_4059_2018[3],
  iro_anaphP_4059_2019[3], iro_anaphP_4059_2020[3],
  iro_anaphP6079[3], iro_anaphP_6079_2017[3], iro_anaphP_6079_2018[3],
  iro_anaphP_6079_2019[3], iro_anaphP_6079_2020[3],
  #### GBS 
    iro_gbs[3], iro_gbs_2017[3], iro_gbs_2018[3], 
  iro_gbs_2019[3], iro_gbs_2020[3], 
  iro_gbs0019[3], iro_gbs_0019_2017[3], iro_gbs_0019_2018[3],
  iro_gbs_0019_2019[3], iro_gbs_0019_2020[3], 
  iro_gbs2039[3], iro_gbs_2039_2017[3], iro_gbs_2039_2018[3],
  iro_gbs_2039_2019[3], iro_gbs_2039_2020[3],
  iro_gbs4059[3], iro_gbs_4059_2017[3], iro_gbs_4059_2018[3],
  iro_gbs_4059_2019[3], iro_gbs_4059_2020[3],
  iro_gbs6079[3], iro_gbs_6079_2017[3], iro_gbs_6079_2018[3],
  iro_gbs_6079_2019[3], iro_gbs_6079_2020[3]
  ))
```

```{r}
Big_table <- data.table(Diagnosis = all_diag, 
                        Age_Bands = all_ages,
                        Year = all_years,
                        Count = count_events,
                        Person_Years = round(pers_year_all, 2),
                        IR_1000_PY = IR_1000PY, 
                        IR_1000_PY_95L = IR_1000PY_95L,
                        IR_1000PY_95H = IR_1000PY_95H)
```


#### Forest Plots {-.tabset} 

##### Total  {-#tot}


```{r}
library(tidyverse)

df <- Big_table 



df_total <- df %>%  
  filter(Year=='Total')

df_total$y_order <- factor(paste(df_total$Age_Bands), levels = unique(paste( df_total$Age_Bands)), ordered = TRUE)
# plot the forest plot
ggplot(df_total, aes(x = IR_1000_PY, y = y_order, size=Person_Years)) +
  geom_point(shape=15) +
  geom_errorbarh(aes(xmin = IR_1000_PY_95L, xmax = IR_1000PY_95H), size = 0.5) +
  facet_grid(Diagnosis ~ ., scales = "free_y", space = "free_y") +
  theme_bw() +
  labs(title = "Forest Plot of Incidence Rates",
       x = "Incidence Rate per 1000 Person-Years",
       y = "Diagnosis and Age Band",
       subtitle = "Data from 2017 to 2020")+
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 160, by = 20)) +
  scale_size_continuous(limits=c(0,2500), breaks = seq(0,2500, by=500))
```

##### 2017 {-#2017}


```{r}

df_2017_P <- df %>%  
  filter(Year=='2017')

df_2017_P$y_order <- factor(paste(df_2017_P$Age_Bands), levels = unique(paste(df_2017_P$Age_Bands)), ordered = TRUE)
# plot the forest plot
ggplot(df_2017_P, aes(x = IR_1000_PY, y = y_order, size=Person_Years)) +
  geom_point(shape=15) +
  geom_errorbarh(aes(xmin = IR_1000_PY_95L, xmax = IR_1000PY_95H), size = 0.5) +
  facet_grid(Diagnosis ~ ., scales = "free_y", space = "free_y") +
  theme_bw() +
  labs(title = "Forest Plot of Incidence Rates",
       x = "Incidence Rate per 1000 Person-Years",
       y = "Diagnosis and Age Band",
       subtitle = "Data from 2017") +
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 160, by = 20)) +
  scale_size_continuous(limits=c(0,2500), breaks = seq(0,2500, by=500))
```


##### 2018 {-#2018}

```{r}

df_2018_P <- df %>%  
  filter(Year=='2018')

df_2018_P$y_order <- factor(paste(df_2018_P$Age_Bands), levels = unique(paste(df_2018_P$Age_Bands)), ordered = TRUE)
# plot the forest plot
ggplot(df_2018_P, aes(x = IR_1000_PY, y = y_order, size = Person_Years)) +
  geom_point(shape=15) +
  geom_errorbarh(aes(xmin = IR_1000_PY_95L, xmax = IR_1000PY_95H), size = 0.5) +
  facet_grid(Diagnosis ~ ., scales = "free_y", space = "free_y") +
  theme_bw() +
  labs(title = "Forest Plot of Incidence Rates",
       x = "Incidence Rate per 1000 Person-Years",
       y = "Diagnosis and Age Band",
       subtitle = "Data from 2018") +
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 160, by = 20)) +
  scale_size_continuous(limits=c(0,2500), breaks = seq(0,2500, by=500))
```

 
##### 2019 {-#2019}

```{r}

df_2019_P <- df %>%  
  filter(Year=='2019')

df_2019_P$y_order <- factor(paste(df_2019_P$Age_Bands), levels = unique(paste(df_2019_P$Age_Bands)), ordered = TRUE)
# plot the forest plot
ggplot(df_2019_P, aes(x = IR_1000_PY, y = y_order, size=Person_Years)) +
  geom_point(shape=15) +
  geom_errorbarh(aes(xmin = IR_1000_PY_95L, xmax = IR_1000PY_95H), size = 0.5) +
  facet_grid(Diagnosis ~ ., scales = "free_y", space = "free_y") +
  theme_bw() +
  labs(title = "Forest Plot of Incidence Rates",
       x = "Incidence Rate per 1000 Person-Years",
       y = "Diagnosis and Age Band",
       subtitle = "Data from 2019")+
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 160, by = 20)) +
  scale_size_continuous(limits=c(0,2500), breaks = seq(0,2500, by=500))
```

##### 2020 {-#2020}

```{r}

df_2020_P <- df %>%  
  filter(Year=='2020')

df_2020_P$y_order <- factor(paste(df_2020_P$Age_Bands), levels = unique(paste(df_2020_P$Age_Bands)), ordered = TRUE)
# plot the forest plot
ggplot(df_2020_P, aes(x = IR_1000_PY, y = y_order, size=Person_Years)) +
  geom_point(shape=15) +
  geom_errorbarh(aes(xmin = IR_1000_PY_95L, xmax = IR_1000PY_95H), size = 0.5) +
  facet_grid(Diagnosis ~ ., scales = "free_y", space = "free_y") +
  theme_bw() +
  labs(title = "Forest Plot of Incidence Rates",
       x = "Incidence Rate per 1000 Person-Years",
       y = "Diagnosis and Age Band",
       subtitle = "Data from 2020")+
  scale_x_continuous(limits = c(0, 150), breaks = seq(0, 160, by = 20)) + 
  scale_size_continuous(limits=c(0,2500), breaks = seq(0,2500, by=500))
```



### Lessons learned {-#less}
Annex 1: quality check findings day 6








## Annex 1 Results Table {-#Tables}

```{r tab2, results = 'hold'}
datatable(Big_table, 
          caption='Incidence Rates with 95% CI',
          class='cell-border stripe', 
          rownames=F, 
          colnames=c('Diagnosis', 'Age Bands', 'Year', 'Amount of Events', 
                     'Person Years', 'IR 1000 Person Years', 'IR 1000 Person Years  95% CI Lower', 'IR 1000 Person Years 95% CI Higher'),
          extensions='Buttons',
          options = list(dom = 'Bfrtip', buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
          ))
```




## Annex 2 Day 6 {-#day6}

```{r fig.align='center', out.width='100%',out.height= 5000,results='asis'}
knitr::include_graphics("day6.pdf")
```



